Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-14 appendix-EPQ.tex
I change proof of EPQ1 and add explanation to it

@@

Let $U$, $V$ and $W$ be the set of operations of $u$, $v$ and $w$, respectively. Let us change $U$, $V$ and $W$ by loops as follows: In the first loop, we start from the first $W$-element of $l'_v$, and let it be $o_h$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Case $1$: If in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and no element in $W$ happens before $o_h$. Then, we erase $o_h$ from $W$ and put it into $U \cdot V$.

\item[-] Case $2$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and there exists $o_w \in W$, such that $o_w <_{\textit{hb}} o_h$. Then, in $l'_v$, we erase all such operations from $U \cdot V$ and put them into $W$, and then stop the process of changing $U$, $V$ and $W$. These operations should satisfy (1) their linearization point is after $o_h$ in $e_{\textit{lp}}$, and (2) the priority of their item is incomparable with $\textit{pri}_x$.

\item[-] Case $3$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is after $\textit{ret}(\textit{rm},x)$. Then, in $l'_v$, we erase all such operations from $U \cdot V$ and put them into $W$, and then stop the process of changing $U$, $V$ and $W$. These operations should satisfy (1) their linearization point is after $o_h$ in $e_{\textit{lp}}$, and (2) the priority of their item is incomparable with $\textit{pri}_x$.
\end{itemize}

Our process proceed, until either all element in $l'_v$ are in new $U$ or new $V$, or case $2$ or case $3$ holds and this process terminates. Let $U'$, $V'$ and $W'$ be the new version of $U$, $V$ and $W$ after the process terminates, respectively. Let $O_+$ be the set of operations that are moved into $U' \cup V'$ in the process, and let $O_-$ be the set of operations that are moved into $W'$ in the process.

Let $l'_{\textit{u'v'}}$ be the projection of $l'$ into $U' \cup V'$, let $O_x$ be the set of $\textit{put}(\_,\textit{pri}_x)$ while the item is not $x$ in $h$. Let $l''_a$ be the longest prefix of $l'_{\textit{u'v'}}$, where linearization of each operation of $l''_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$. Let $l''_d$ be the projection of $l'$ into operations of $O_x$ that are not in $l''_a$. Let $l''_1 = l''_a \cdot l''_d$. Let $l''_2$ be the projection of $l'$ into operations of $l'_{\textit{u'v'}}$ that are not in $l''_1$. Let $l''_3$ be the projection of $l'$ into operations of $l'_{\textit{u'v'}}$ that are not in $l''_1$ and $l''_2$. Let $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$.

@@

\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,\textit{pri}_x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_1 \in O_a$: This is impossible, since the linearization point of operations in $O_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$.

    \item[-] If $o_1 \in O_d$: Then $l$ contradicts with happen before relation of $h$.
    \end{itemize}

@@

    \item[-] If $o_2 \in O_d$: According to $l$, we can see that $\textit{put}(x,\textit{pri}_x)$ does not happen before any operation in $O_x$. Then we can see that the linearization point of $o_1$ is before $\textit{ret}(\textit{put},b)$ and $o_1 \in O_a$. This violates that $o_1 \in l''_2$.
    \end{itemize}

@@

    \item[-] If $o_2 \in O_+$: This is impossible, since the linearization point of operations in $O_+$ is before $\textit{ret}(\textit{rm},b)$ in $e_{\textit{lp}}$.
    \end{itemize}

@@

    \item[-] If $o_1 \in W \wedge o_2 \in O_+$: This is impossible, since according to construction of $O_+$, we can see that $o_2 \in O_-$.

    \item[-] If $o_1 \in O_- \wedge o_2 \in U \cup V$:
        \begin{itemize}
        \setlength{\itemsep}{0.5pt}
        \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$. Therefore, we can see that $o_w <_{\textit{hb}} o_2$, and then $l$ contradicts with happen before relation of $h$.

@@

    \item[-] If $o_1 \in O_-$:
         \begin{itemize}
         \setlength{\itemsep}{0.5pt}
         \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(\textit{put},x,\textit{pri}_x)$. Therefore, we can see that $o_w <_{\textit{hb}} \textit{put}(x,\textit{pri}_x)$, and then $l$ contradicts with happen before relation of $h$.

@@

    \item[-] If $o_1 \in O_-$:
         \begin{itemize}
         \setlength{\itemsep}{0.5pt}
         \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$.

@@

It remains to prove that $l'' \in \textit{EPQ}$. Let $O_c$ be the set of operations in $h$, whose items have priority that are comparable with $\textit{pri}_x$, and let $O_i$ be the set of operations in $h$, whose items have priority that are incomparable with $\textit{pri}_x$. Then we obtain that $l'' \in \textit{EPQ}$ as follows:

@@

\item[-] $l'_{\textit{u'v'}}$ can be obtained from $l'_v$ as follows: From some time point $t$, discard all the $O_i$ operations after $t$ in $l'_v$. Since $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $\forall p_i \in O_i$ and $\forall p_c \in O_c$, either $p_c <_{\mathbb{P}} p_i$, or $p_c$ is incomparable with $p_i$. Therefore, this process does not influence correctness of $l'_{\textit{u'v'}}$, and then $l'_{\textit{u'v'}} \in \textit{EPQ}$.

\item[-] $l''_1 \cdot l''_2$ can be obtained from $l'_{\textit{u'v'}}$ as follows: Execute until reaching some time point $t$, then first execute all $O_x$ operations after $t$, and then execute remanning operations. Since $O_x$ only contains $\textit{put}(\_,\textit{pri}_x)$ and $\textit{pri}_x$ is one of maximal priority in $h$, we can see that $l''_1 \cdot l''_2 \in \textit{EPQ}$.

\item[-] Let $l''_f$ be the projection of $l'$ into $O_-$, and let $l''_g$ be the projection of $l'$ into operations not in $l''_1$, $l''_2$ or $l''_f$. It is easy to see that $l''_3 = l''_f \cdot l''_g$.

>>>>

Before we prove the step-by-step linearizability of $\textit{MS}(\textit{EPQ}_1)$ and $\textit{MS}(\textit{EPQ}_2)$, we introduce several lemmas, which are used to ensure some sub-sequences of $\textit{EPQ}$ still belongs to $\textit{EPQ}$. Given a data-differentiated sequence $l$ and one of its maximal priority $\textit{pri}$, let $O_c(l,\textit{pri})$ and $O_i(l,\textit{pri})$ be the set of operations with priorities comparable with $\textit{pri}$ and incomparable with $\textit{pri}$ in $l$, respectively. Similarly we can define $D_c(l,\textit{pri})$ and $D_i(l,\textit{pri})$ for items instead of operations. %Let $O_x(l)$ be the set of $\textit{put}(y,\textit{pri})$ with $y \neq x$ in $l$, where $\textit{pri}$ is the priority of $x$ in $l$.
We can see that each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in  $O_c(l,\textit{pri})$.

The following lemma shows that if a new sequence is generated by erasing some operations in $O_c(l,\textit{pri})$ while keeping the remaining $O_c(l,\textit{pri})$ sub-sequences in $\textit{EPQ}$, then this new sequence is still in $\textit{EPQ}$. Note that this is different from projection on items.

\begin{restatable}{lemma}{EraseOcStillinEPQ}
\label{lemma:erase Oc still in EPQ}
Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$. Let $l'$ be generated from $l$ by discarding some operations in $O_c(l,\textit{pri})$, and $l' \vert_{ O_c(l,\textit{pri}) } \in \textit{EPQ}$. Then, $l' \in \textit{EPQ}$.
\end{restatable}

\begin {proof}
Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$. Assume that $l'$ is generated from $l$ by discarding $o_{\textit{ind1}},\ldots,o_{\textit{indn}}$. Let $D$ be the set such that $D$ contains $a$, if $\textit{put}(a,\_)$ is in $o_{\textit{ind1}},\ldots,o_{\textit{indn}}$. For each $i$, let $q'_i$ be generated from $q_i$ by erasing items in $D$.

For each $q'_j$ with $j \neq \textit{ind}\_\textit{-1}$, if $o_{\textit{j+1}}$ is $\textit{put}$, then it is obvious that $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$. Else, assume $o_{\textit{j+1}} = \textit{rm}(a)$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{rm}(a) \in O_c(l,\textit{pri})$: By assumption, $l' \vert_{ O_c(l,\textit{pri}) } \in \textit{EPQ}$. Therefore, $a$ is in $q'_j$ and is the should-be-removed item in $O_c(l,\textit{pri})$. Since each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in $O_c(l,\textit{pri})$, we can removed $a_c$ from $q'_j$, and then $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.

\item[-] If $\textit{rm}(a) \in O_i(l,\textit{pri})$: By assumption we know that $q_j \xrightarrow{\textit{rm}(a)} q_{\textit{j+1}}$. Since $q'_j$ contains the same $D_i(l,\textit{pri})$ items as $q_j$ and $q'_j$ contains less $D_c(l,\textit{pri})$ items than $q_j$, we can see that $q'_j \xrightarrow{o_{\textit{j+1}}} q'_{\textit{j+1}}$.
\end{itemize}

For each $q'_{\textit{indj-1}}$, it is easy to see that $q'_{\textit{indj-1}} = q'_{\textit{indj}}$. Therefore, we can see that $l' = o_1 \cdot \ldots o_{\textit{ind1-1}} \cdot o_{\textit{ind1+1}} \cdot \ldots \in \textit{EPQ}$. \qed
\end {proof}

The following lemma shows that if a new sequence is generated by from some time point, erasing operations in $O_i(l,\textit{pri})$, then this new sequence is still in $\textit{EPQ}$.

\begin{restatable}{lemma}{EraseOiFromSomeTimePointStillinEPQ}
\label{lemma:erase Oi from some time point still in EPQ}
Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$. Let $l'$ be generated from $l$ by discarding operations in $O_i(l,\textit{pri})$ from some time point, then, $l' \in \textit{EPQ}$.
\end{restatable}

\begin {proof}
Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$. Assume that $l'$ is generated from $l$ by discarding all operations $o_i$ if (1) $o_i \in O_i(l,\textit{pri})$ and (2) $i \geq k$ for a specific index $k$. Let $D$ be a set such that $a \in D$, if $\textit{put}(a,\_)$ is in $l$ and not in $l'$. For each $0 \leq i \leq m$, let $q'_i$ be generated from $q_i$ by erasing items in $D$.

Let $l'=o'_1 \cdot \ldots \cdot o'_n$, and let $f$ be a function, such that $f(i)=j$, if $o'_i = o_j$.

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] We can see that $f$ maps each $0 \leq i \leq \textit{k-1}$ into $i$, and $q'_0 \xrightarrow{o_1} q'_1 \ldots \xrightarrow{o_{\textit{k-1}}} q'_{\textit{k-1}}$.

\item[-] It is easy to see that $q'_{\textit{k-1}} = q'_{f(k)-1}$, and for each $i>k$, $q'_{f(k)} = q'_{f(k+1)-1}$.

\item[-] If $o_{f(k)}$ is a $\textit{put}$ operation, then it is obvious that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$. Else, if $o_{f(k)} = \textit{rm}(a)$, we can see $a$ is in $q'_{f(k)-1}$, and since (1) $q'_{f(k)-1}$ contains the same $D_c(l,\textit{pri})$ items as $q_{f(k)-1}$ and $q'_{f(k)-1}$ contains less $D_i(l,\textit{pri})$ items than $q_{f(k)-1}$, and (2) each priority of items in $O_i(l,\textit{pri})$ is either larger or incomparable with priorities of items in  $O_c(l,\textit{pri})$, we can see that $q'_{f(k)-1} \xrightarrow{o_k} q'_{f(k)}$. Similarly we can prove the case of $o_{f(j)}$ with $j > k$.
\end{itemize}

The following lemma shows that we can make $\textit{put}$ with maximal priority to happen earlier.

\begin{restatable}{lemma}{MakePutWithMaxPriorityHappenEarlier}
\label{lemma:make put with maximal priority happen earlier}
Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$. Let $l=l_1 \cdot l_2$. Let $l_3$ be the projection of $l_2$ into $\{ \textit{put}(\_,\textit{pri}) \}$, and $l_4$ be the projection of $l_2$ into other operations. Then, $l'=l_1 \cdot l_3 \cdot l_4 \in \textit{EPQ}$.
\end{restatable}

\begin {proof}
Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$. Let $l_1 = o_1 \cdot \ldots \cdot o_n$, let $l_2 = o_{\textit{n+1}} \cdot \ldots \cdot o_m$, let $l_3 = o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$, let $l_4 = o'_{\textit{k+1}} \cdot \ldots \cdot o'_m$. Let $f$ be a function, such that $f(i)=j$, if $o'_i = o_j$.

Let $q'_i$ be constructed as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] For $0 \leq i \leq n$, let $q'_i = q_i$.

\item[-] For $\textit{n+1} \leq i \leq k$, let $q'_i$ be obtained from $q_n$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ with priority $\textit{pri}$ and in the same order.

\item[-] For $\textit{k+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(i)}$. The order of adding them is the same as $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$.
\end{itemize}

Then, our proof proceeds as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] It is obvious that $q'_0 \xrightarrow{o_1} q'_1 \ldots \xrightarrow{o_n} q'_n$ and $q'_n \xrightarrow{o_{\textit{n+1}}} q'_{\textit{n+1}} \ldots \xrightarrow{o_k} q'_k$.

\item[-] For $q'_{\textit{k+1}}$: We already know that $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, and it is easy to see that $q_{f(k+1)-1}$ is obtained from $q_n$ by adding items in $o_{\textit{n+1}} \cdot \ldots \cdot o_{f(k+1)-1}$.

    We can see that $q'_k$ is obtained from $q_n$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$, and $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+1)}$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_{f(k+1)}$ is an operation of non-$\textit{pri}$ items, then $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$. Since non-$\textit{pri}$ priority is either smaller or incomparable with $\textit{pri}$, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.

    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some item $a$ with priority $\textit{pri}$. We can see that $q'_{k+1}$ is obtained from $q_{f(k+1)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$. Since $q_{f(k+1)-1} \xrightarrow{o_{f(k+1)}} q_{f(k+1)}$, in $q_{f(k+1)-1}$( and also in $q'_k$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$. Therefore, we can see that $q'_k \xrightarrow{o_{f(k+1)}} q'_{\textit{k+1}}$.
    \end{itemize}

\item[-] For $q'_{\textit{k+i}}$ with $i>1$: We already know that $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, and $q_{f(k+i)-1}$ is obtained from $q_{f(k+i-1)}$ by adding items in $o_{f(k+i-1)+1} \cdot \ldots \cdot o_{f(k+i)-1}$.

    We can see that $q'_{\textit{k+i}-1}$ (resp., $q'_{\textit{k+i}}$) is obtained from $q_{f(k+i-1)}$ (resp., $q_{f(k+i)}$) by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$ (resp., $o_1 \cdot \ldots \cdot o_{f(k+i)}$).

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_{f(k+i)}$ is an operation of non-$\textit{pri}$ items, then $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding items which are (1) with priority $\textit{pri}$, (2) in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_i$ and not removed by $o_1 \cdot \ldots \cdot o_{f(k+i-1)}$. Since non-$\textit{pri}$ priority is either smaller or incomparable with $\textit{pri}$, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.

    \item[-] Otherwise, it is only possible that $o_{f(k+i)} = \textit{rm}(a)$ for some item $a$ with priority $\textit{pri}$. We can see that $q'_{k+i}$ is obtained from $q_{f(k+i)}$ by adding items in $o'_{\textit{n+1}} \cdot \ldots \cdot o'_k$ and then remove $a$. Since $q_{f(k+i)-1} \xrightarrow{o_{f(k+i)}} q_{f(k+i)}$, in $q_{f(k+i)-1}$( and also in $q'_{\textit{k+i-1}}$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$. Therefore, we can see that $q'_{\textit{k+i-1}} \xrightarrow{o_{f(k+i)}} q'_{\textit{k+i}}$.
    \end{itemize}
\end{itemize}

The following lemma shows that if a new sequence is generated by make some $O_i(l,\textit{pri})$ behaviors to happen earlier, then this new sequence is still in $\textit{EPQ}$.

\begin{restatable}{lemma}{MakeOiHappenEarlierStillinEPQ}
\label{lemma:make Oi happen earlier still in EPQ}
Given a data-differentiated sequential execution $l \in \textit{EPQ}$ and a maximal priority $\textit{pri}$ in $l$, where $l$ does not contain $\textit{rm}(\textit{empty})$. Let $l \vert_{ O_i(l,\textit{pri}) } = l_1 \cdot l_2$, let $l' = l_1 \cdot l_3$, where $l_3$ is the projection of $l$ into non-$l_1$ operations. Then, $l' \in \textit{EPQ}$.
\end{restatable}

\begin {proof}
Let $l=o_1 \cdot \ldots \cdot o_m$, and $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$. Let $l_1 = o'_1 \cdot \ldots \cdot o'_n$, let $l_3 = o'_{\textit{n+1}} \cdot \ldots \cdot o'_m$. Let $f$ be a function, such that $f(i)=j$, if $o'_i = o_j$. Let $D$ be the se of items which are added and not removed in $o'_1 \cdot \ldots \cdot o'_n$.

Let $q'_i$ be constructed as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] It is easy to see that $l_1 \in \textit{EPQ}$, and let $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_1$ in $\textit{LTS}_e$.

\item[-] For $\textit{n+1} \leq i \leq m$, let $q'_i$ be obtained from $q_{f(i)}$ by adding items in $D$. The order of adding them is the same as $o'_1 \cdot \ldots \cdot o'_n$.
\end{itemize}

Then, our proof proceeds as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] We already know that $q'_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$.

\item[-] For $q'_{\textit{n+i}}$: We already know that $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, and it is easy to see that $q_{f(n+i)-1}$ is obtained from $q_{f(n+i-1)}$ by adding $D$-items in $o_{f(n+i-1)+1} \cdot \ldots \cdot q_{f(n+i)-1}$.

     We can see that $q'_{\textit{n+i}-1}$ (resp., $q'_{\textit{n+i}}$) is obtained from $q_{f(n+i-1)}$ (resp., $q_{f(n+i)}$) by adding remanning items in $D$.

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_{f(n+1)}$ is an operation of $O_c(l,\textit{pri})$ items, since priority in $O_i(l,\textit{pri})$ is either larger or incomparable with priority in $O_c(l,\textit{pri})$, we can see that $q'_n \xrightarrow{o_{f(n+1)}} q'_{\textit{n+1}}$.

    \item[-] Otherwise, it is only possible that $o_{f(k+1)} = \textit{rm}(a)$ for some item $a$ in $O_c(l,\textit{pri})$. Since $q_{f(n+i)-1} \xrightarrow{o_{f(n+i)}} q_{f(n+i)}$, in $q_{f(n+i)-1}$( and also in $q'_{\textit{n+i-1}}$), there is no item with priority less than $\textit{pri}$, and $a$ is the first-input item of priority $\textit{pri}$. Therefore, we can see that $q'_{\textit{n+i-1}} \xrightarrow{o_{f(n+i)}} q'_{\textit{n+i}}$.
    \end{itemize}
\end{itemize}

The following lemma shows that if a new sequence is generated by replacing a prefix with another one which make the priority queue has same content, then this new sequence is still in $\textit{EPQ}$.

\begin{restatable}{lemma}{ReplaceEquivalentPrefixStillinEPQ}
\label{lemma:replace equivalent prefix still in EPQ}
Given a data-differentiated sequential execution $l \in \textit{EPQ}$. Let $l=l_1 \cdot l_2$. Given $l_3 \in \textit{EPQ}$. Assume that the priority queue has same content after executing $l_1$ and $l_3$. Let $l' = l_3 \cdot l_2$. Then, $l' \in \textit{EPQ}$.
\end{restatable}

\begin {proof}
Let $l=o_1 \cdot \ldots \cdot o_m$ and let $q_0 \xrightarrow{o_1} q_1 \ldots \xrightarrow{o_m} q_m$ be the path of $l$ in $\textit{LTS}_e$. Let $l_1 = o_1 \cdot \ldots \cdot l_k$, $l_3 = o'_1 \cdot \ldots \cdot o'_n$ and let $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n$ be the path of $l_3$ in $\textit{LTS}_e$.

By assumption we know that $q_k = q'_n$. Then it is not hard to see that $q_0 \xrightarrow{o'_1} q'_1 \ldots \xrightarrow{o'_n} q'_n \xrightarrow{o_{\textit{k+1}}} q_{\textit{k+1}} \ldots \xrightarrow{o_m} q_m$ is a path in $\textit{LTS}_e$, and then $l' \in \textit{EPQ}_s$. By Lemma \ref{lemma:EPQ rules and semantics}, we know that $l' \in \textit{EPQ}$. \qed
\end {proof}

With the help of Lemma \ref{lemma:erase Oc still in EPQ}, Lemma \ref{lemma:erase Oi from some time point still in EPQ}, Lemma \ref{lemma:make put with maximal priority happen earlier}, Lemma \ref{lemma:make Oi happen earlier still in EPQ} and Lemma \ref{lemma:replace equivalent prefix still in EPQ}, we can now prove that $\textit{EPQ}_1$ is step-by-step linearizability.

@@

Let $U$, $V$ and $W$ be the set of operations of $u$, $v$ and $w$, respectively. Let us change $O_i(l,\textit{pri}_x)$ elements in $U$, $V$ and $W$, while keep $O_c(l,\textit{pri}_x)$ elements unchanged. We proceed by several loops: In the first loop, we start from the first $O_i(l,\textit{pri}_x)$-element of $l'_v$, and let it be $o_h$,

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] Case $1$: If in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and no $O_c(l,\textit{pri}_x)$-element in $W$ happens before $o_h$. Then, $o_h$ is in the new version of $U \cup V$.

\item[-] Case $2$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is before $\textit{ret}(\textit{rm},x)$, and there exists $O_c(l,\textit{pri}_x)$-element $o_w$ in $W$, such that $o_w <_{\textit{hb}} o_h$. Then, in $l'_v$, we put $o_h$ and all $O_i(l,\textit{pri}_x)$-element whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.

\item[-] Case $3$: Else, if in $e_{\textit{lp}}$, the linearization point of $o_h$ is after $\textit{ret}(\textit{rm},x)$. Then, in $l'_v$, we put $o_h$ and all $O_i(l,\textit{pri}_x)$-element whose linearization points is after the linearization point of $o_h$ into new version of $W$, and then stop the process of changing $U$, $V$ and $W$.
\end{itemize}

In the next loop, we consider the second $O_i(l,\textit{pri}_x)$-element of $l'_v$, and so on. Our process proceed, until either all element in $l'_v$ are in new version of $U \cup V$, or case $2$ or case $3$ happens and this process terminates. Let $U' \cup V'$ and $W'$ be the new version of $U \cup V$ and $W$ after the process terminates, respectively. Let $O_+$ be the set of operations that are moved into $U' \cup V'$ in the process, and let $O_-$ be the set of operations that are moved into $W'$ in the process.

Let $l'_{\textit{u'v'}}$ be the projection of $l'$ into $U' \cup V'$, let $O_x$ be the set of $\textit{put}(\_,\textit{pri}_x)$ while the item is not $x$ in $h$. Let $l''_a$ be the longest prefix of $l'_{\textit{u'v'}}$, where linearization of each operation of $l''_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$. Let $l''_d$ be the projection of $l'$ into operations of $O_x$ that are not in $l''_a$. Let $l''_1 = l''_a \cdot l''_d$. Let $l''_2$ be the projection of $l'$ into operations of $l'_{\textit{u'v'}}$ that are not in $l''_1$. Let $l''_3$ be the projection of $l'$ into operations which are not in $l'_{\textit{u'v'}}$. Let $l'' = l''_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3$.

@@

\item[-] If $o_2 \in l''_1 \wedge o_1 = \textit{put}(x,\textit{pri}_x)$:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_2 \in O_a$: This is impossible, since the linearization point of operations in $O_a$ is before $\textit{ret}(\textit{put},b)$ in $e_{\textit{lp}}$.

    \item[-] If $o_2 \in O_d$: Then $l$ contradicts with happen before relation of $h$.
    \end{itemize}

@@

    \item[-] If $o_2 \in O_d$: According to $l$, we can see that $\textit{put}(x,\textit{pri}_x)$ does not happen before any operation in $O_x$. Then we can see that the linearization point of $o_1$ is before $\textit{ret}(\textit{put},x)$ and $o_1 \in O_a$. This violates that $o_1 \in l''_2$.
    \end{itemize}

@@

    \item[-] If $o_2 \in O_+$: This is impossible, since the linearization point of operations in $O_+$ is before $\textit{ret}(\textit{rm},x)$ in $e_{\textit{lp}}$.
    \end{itemize}

@@

    \item[-] If $o_1 \in W \wedge o_2 \in O_+$:

    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] If $o_1 \in O_i(l,\textit{pri}_x)$: Then according to the construction process of $U' \cup V'$ and $W'$, we can see that $o_1 \in O_+$ and then $o_1 \in U' \cup V'$, which contradicts that $o_1 \in l''_3$.

    \item[-] If $o_1 \in O_c(l,\textit{pri}_x)$: Then according to the construction process of $U' \cup V'$ and $W'$, we can see that $o_2 \in O_-$, which contradicts that $o_2 \in O_+$.
    \end{itemize}

    \item[-] If $o_1 \in O_- \wedge o_2 \in U \cup V$:
        \begin{itemize}
        \setlength{\itemsep}{0.5pt}
        \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(o_2)$. Therefore, we can see that $o_w <_{\textit{hb}} o_2$, and then $l$ contradicts with happen before relation of $h$.

@@

    \item[-] If $o_1 \in O_-$:
         \begin{itemize}
         \setlength{\itemsep}{0.5pt}
         \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$, and $\textit{ret}(o_1)$ is before $\textit{cal}(\textit{put},x,\textit{pri}_x)$. Therefore, we can see that $o_w <_{\textit{hb}} \textit{put}(x,\textit{pri}_x)$, and then $l$ contradicts with happen before relation of $h$.

@@

    \item[-] If $o_1 \in O_-$:
         \begin{itemize}
         \setlength{\itemsep}{0.5pt}
         \item[-] If the reason of $o_1 \in O_-$ is case $2$: Let $o_h$ be as in case $2$. Then there exists $O_c(l,\textit{pri}_x)$-element $o_w \in W$, and in $e_{\textit{lp}}$, $\textit{ret}(o_w)$ is before $\textit{cal}(o_h)$, the linearization point of $o_h$ is before the linearization point of $o_1$.

@@

It remains to prove that $l'' \in \textit{EPQ}$. The process for proving $l'' \in \textit{EPQ}$ is as follows:

@@

\item[-] $l'_{\textit{u'v'}}$ can be obtained from $l'_v$ as follows:
    \begin{itemize}
    \setlength{\itemsep}{0.5pt}
    \item[-] Discard $O_c(l,\textit{pri}_x)$-element that are in $W$ and keep $O_c(l,\textit{pri}_x)$-element in $U \cup V$ unchanged.

    \item[-] From some time point, discard all the $O_i(l,\textit{pri}_x)$ operations after this time point.
    \end{itemize}

From $\textit{matched-C}(u \cdot v,\textit{pri}_x)$, we can see that $O_c(l,\textit{pri}_x)$-element in $U \cdot V$ is matched, and then it is easy to see that the projection of $l'_v$ in to $O_c(l,\textit{pri}_x)$-element in $U \cdot V$ is still in $\textit{EPQ}$. By Lemma \ref{lemma:erase Oc still in EPQ} and Lemma \ref{lemma:erase Oi from some time point still in EPQ}, we can see that $l'_{\textit{u'v'}} \in \textit{EPQ}$.
